Two-Dimensional Hydrodynamics of a Janus Particle Vesicle

### 2022-02-16 doc/reviews/ResponseToReviewer2.tex

Finally we provide a critical capillary number for rupture of a JP vesicle under a linear shear flow. The rupture of a lipid vesicle by a linear shear flow has been reported in experiments.  Our prediction of a critical capillary number and how it scales with various physical parameters of the individual Janus particle is, to our knowledge, not available in the literature. We hope that our results can motivate experimentalist (such as Charles Schroeder who has studied extreme deformation of a vesicle due to an imposed flow) and other theoreticians to study the dynamics of bilayer structures under extreme flowing conditions.

%  \begin{quotation}
%    \noindent
%    Can put verbatim change in this environment.
%  \end{quotation}
\end{itemize}

\noindent
\comment{The abbreviation HAP is not defined in the text. I guess it stands for hydrophobic attraction potential.
}
\begin{itemize}
  \item Response: Thank you for noticing this oversight. 
  \item Change: The definition of HAP has been added at its first occurrence in Section 3.1. 
\end{itemize}

\noindent
\comment{$u$ is used for both the velocity and the solutions of the
Laplace equation. Please use another variable for the scalar function
$u$.}
\begin{itemize}
\item Response: The velocity ${\bf u}({\bf x})$ introduced in the mobility problem (eqn. (2.1)--(2.3)) is in bold face and of a vector form. The scalar solution function $u$ is the solution of the screened Laplace equation boundary value problem (eqn. (2.7)--(2.9)).
  We can see, however, how this might result in confusion. 
\item Change: In response to this point, we changed the scalar function $u$ to $\eta$ throughout the manuscript.
  The notation $\eta$ was used by Stjepan Marcelja to first pose the HAP formulation
  (Croatica Chemica Acta, 1977).
\end{itemize}

@@

\noindent
\comment{$\rho$ used in the HAP is not the same as the one used in the
mobility boundary integral equation. One is a decay length of the
attraction potential, while the second is the Euclidean distance between
two JPs.}
\begin{itemize}
  \item Response: We had used $\rho$ to denote the distance between source and
    target points in one of the layer potentials. This has been removed.

  \item Change: We now use $\rho$ to denote the HAP decay length and $\rho_0$ to
    denote the cutoff for the repulsion force.
\end{itemize}

\noindent
\comment{$\nu$ is used on multiple instances, sometimes as a normal
vector (see line 97 for example), and other times as some scalar
function (see line 203).}
\begin{itemize}
  \item Response: We thank the referee for this oversight.
  \item Change: We now reserve $\boldsymbol{\nu}$ for the normal vector and
    $w_i$ denote the scalar what we previously defined to be $\nu_i$.
\end{itemize}

\noindent
\comment{What is the bending modulus here? I guess that it is not a
direct input parameter in this model but rather a quantity that can be
deduced from other input parameters. If I am right then, what are these
parameters? And could you tune them if needed to simulate the effect of
the membrane stiffness? This can be important in some applications such
as rigidity-based microfluidic sorting devices.}
\begin{itemize}
\item Response: The Reviewer is correct that realistic bending moduli
  and other elastic parameters can be deduced from simulation data.
  At this point, however, we do not have an analytical connection between
  the input and moduli and inputs, other than some scaling relationship.
  Please see the next comment, though, regarding the bending and stretching
  moduli.
\item Change: In the beginning of Section 4.1,
  we now point out that such an analysis was part of our previous work.  
\end{itemize}

\noindent
\comment{To compare the results with the existing literature, I would
suggest introducing the capillary number for vesicles as defined in
\url{https://doi.org/10.1016/j.crhy.2009.10.001}.}
\begin{itemize}
\item Response:  We thank the Reviewer for pointing out this interesting connection. 
\item Change:  Reviewer 1 requested a parametric study of membrane rupture which
  we now supply in Section 4.2.3. There we take up the present Reviewer's
  suggestion and show how the critical rupture shear rate ties in with the capillary
  number, and then use the capillary number to derive an effective bending rigidity. 
\end{itemize}

\noindent
\comment{For both examples with the linear shear flow and the Poiseuille
flow, the area of the vesicle is not conserved and the relative error
with respect to the stress-free area increases with the increase of the
external stresses. It can reach roughly 7-8\% for the higher shear rate
in the shear flow example. This is quite significant. Do you have a way
to reduce this error to below 1\%? Note that in other numerical models
for vesicles (e.g.~\url{https://doi.org/10.1016/j.jcp.2008.11.036}),
both constant area and perimeter constraints are fulfilled even when the
vesicle is more deformed than the cases reported here.}
\begin{itemize}
\item Response: We thank the referee for this comment. For impermeable vesicles, the referee is correct about the numerical errors reported in the literature. However, the JP membrane is permeable and in our simulations we found that the Janus vesicle needs time so the Janus particles in the membrane can re-arrange themselves. During this process of rearrangement the JP bilayer membrane behaves as a permeable membrane as in Quaife, Gannon and Young (PRF 2021). For a JP vesicle in a linear shear flow with a shear rate of dimensionless shear $\chi = 0.002$ the enclosed area reaches a nearly constant value after $1$ $\mu$s. We would like to point out that this is not a numerical error, but a consequence of Janus particle hydrodynamics that gives rise to membrane permeability.

\item Change: In response to Reviewer 1, we have rewritten Section 4.3 from scratch.
  This section now performs a parametric study of membrane permeability.
\end{itemize}

\noindent
\comment{In some interesting applications, the fluid encapsulated by the
bilayer membrane has a viscosity which is different from the suspending
fluid. How difficult it is to add this feature to your model?}
\begin{itemize}
\item Change:
  We now mention in the conclusion that our current numerical method does not support a viscosity contrast in the system 
or the setup similar to the ones in (Veerapaneni {\sl et al.}, {\bf 2011}, Phys. Rev. Lett.),
(Vlahovska and Gracia, {\bf 2007}, Phys. Rev. E), and
(Kaoui and Harting, {\bf 2016}, Rheol Acta). 
As described in the manuscript, the JP vesicle is permeable and the mixture of two different
fluids has to be analyzed within the framework, which would introduce a contact line between fluid 1- JP and fluid 2 -JP boundaries on the Janus particle surface. It is an interesting problem and we plan to study this in the future after we establish some basic knowledge of JP vesicle dynamics in a simpler configuration here.

### 2022-02-16 doc/main_jfm.tex
main and reviewer 2

\subsection{Model Parameters}
\cite{Fu20} studied physical quantities for static JP vesicles. 
In the present work, we study JP vesicles in various background flows.
We also perform parametric studies of effective properties of the JP vesicles 
as a function of particle size and interaction strength.
An analysis of the relationship between particle shape on effective properties is beyond 
the scope of the current paper.

@@

\ReviewerOne{An emperical numerical study found that for the baseline parameter set,
the dimensionless, critical shear rate is $\chi_*=0.0655$ so that the membrane ruptured whenever $\chi > \chi_*$
  and did not rupture whenever $\chi < \chi_*.$ 
  Supplementary Materials Movie 4 shows the membrane rupture at the critical shear rate.
  To further quantify the competition between shear and attraction,
  we found that a 5 \% relative increase in the parameters $\gamma$ and $M$ 
  resulted in a  4.24 \% relative increase in $\chi_*$, from $0.0655$ to $0.0688$.
  A 5 \% decrease in the particle radius $c$ gave a -9.55 \% decrease in $\chi_*$
  and a 10 \% increase in resting length $L_0$ gave a 20.61 \% decrease in critical shear rate $\chi_*$.
  These relative changes suggest that there is dimensionless number $K$ given by 
  \begin{equation}
    \label{eq:CSR}
  \mu \dot\gamma_* = K \frac{\gamma c^2}{\rho L^2}
  \end{equation}
  that relates hydrophobic attraction to the dimensional critical shear rate $\dot \gamma_* = \chi_* \text{ ns}^{-1}$.
  Assuming a characteristic velocity $L\dot \gamma$, the number $K$ relates to the vesicle's capillary number
  $\text{Ca} = K c^2/(\rho L)$. Here $\text{Ca}$ would be approximately $1.7$
  so that surface forces are comparable to viscous forces.  
  There is a related \emph{elastic} capillary number
  $\tilde{\text{Ca}} = \mu \dot{\gamma} R_0^3/K_B$
  used to characterize rheological properties of lipid bilayer membranes
  where $R_0$ is vesicle radius $K_B$ is bending rigidity \citep{VLAHOVSKA2009775, keller_skalak_1982}. 
  Setting $\text{Ca}$ and $\tilde{\text{Ca}}$ equal, we can derive
  \[
  K_B = \frac{\mu \dot \gamma_* R_0^3}{\text{Ca}}
  \]
  which yields $K_B \approx 60$ pN $\text{nm}^{-1}$, which is a reasonable
  value for the bending modulus of a lipid bilayer, for example \citep{NaTr00}.}

@@

\subsection{Stretching and Permeability}
\label{sec:ves_in_quiescent}

@@

In membrane continuum  mechanics, changes in membrane surface area $A$ give rise to a tension
$\tilde \sigma = K_A(A/A_0 - 1)$ where $A_0$ is the reference surface area \cite{NaTr00}. 
For a two-dimensional membrane, the tension is 
\begin{align}
\label{eq:stretch}
\tilde \sigma = K_A\left(L/L_0 - 1 \right),
\end{align}
where $L$ and $L_0$ are arc length and resting length,
respectively. 
Membrane tension gives rise to 
a Laplace pressure 
\begin{align}
\label{eq:LaplacePressure}
\Delta p = \tilde \sigma/R,
\end{align}
where $\Delta p$ is the jump in far-field pressure to internal pressure and
$R = L/2\pi$.
In order to relieve Laplace pressure, 
water flows across the membrane. The rate of efflux is governed by  
\begin{align}
  \label{eq:perm} 
  \frac{dA}{dt} = - P\Delta p,
\end{align}
where $P = k/(\mu \Delta x)$ is a constant, $k$ is hydraulic permeability, 
$\mu$ is the solvent viscosity, and $\Delta x$ is the 
membrane thickness, about 5 nm
(\cite{chabanon2017, qua-gan-you2021}).

@@

Our future goals include extending the current framework to a
three-dimensional JP vesicle system. This will require additional
algorithmic implementation including a fast summation method such as the
fast multipole method. Another research direction is to include the
fluctuating hydrodynamics for Brownian suspensions~(\cite{Bao2018}),
%introduce thermal fluctuations into the model. 
%This will result in Janus particles that include Brownian motion, 
and this is critical to understand membrane diffusion. Finally, a more
physical boundary conditions for the HAP model will allow us to draw
comparisons between computational and laboratory experiments.

>>>>

\subsection{Model Parameters}
\cite{Fu20} studied physical quantities for static JP vesicles
\ReviewerTwo{and empitically deduced realistic bending moduli, stretching moduli,
  and tilt moduli under reasonable assumptions for the input parameters.}
In the present work, we study JP vesicles in various background flows.
We also perform parametric studies of effective properties of the JP vesicles 
as a function of particle size and interaction strength.

@@

\ReviewerOne{An emperical numerical study found that for the baseline parameter set,
the dimensionless, critical shear rate is $\chi_*=0.0655$ so that the membrane ruptured whenever $\chi > \chi_*$
  and did not rupture whenever $\chi < \chi_*.$ 
  Supplementary Materials Movie 4 shows the membrane rupture at the critical shear rate.
  To further quantify the competition between shear and attraction,
  we found that a 5 \% relative increase in the parameters $\gamma$ and $M$ 
  resulted in a  4.24 \% relative increase in $\chi_*$, from $0.0655$ to $0.0688$.
  A 5 \% decrease in the particle radius $c$, screening length $\rho$ gave a $-9.55$ \% decrease, $5.0 \%$ increase
  respectively in $\chi_*$.
  Finally, a 10 \% increase in resting length $L_0$ gave a 20.61 \% decrease in critical shear rate $\chi_*$.}  
  \ReviewerOne{These relative changes suggest that there is dimensionless number $K$ given by 
  \begin{equation}
    \label{eq:CSR}
  \mu \dot\gamma_* = K \frac{\gamma c^2}{\rho L^2}
  \end{equation}
  that relates hydrophobic attraction to the dimensional critical shear rate $\dot \gamma_* = \chi_* \text{ ns}^{-1}$.
  Assuming a characteristic velocity $L\dot \gamma$, the number $K$ relates to the vesicle's capillary number
  $\text{Ca} = K c^2/(\rho L)$. Based on the above data, $\text{Ca}$ would be approximately $1.7$
  so that surface forces are comparable to viscous forces.  
  There is a related \emph{elastic} capillary number
  $\tilde{\text{Ca}} = \mu \dot{\gamma} R_0^3/K_B$
  used to characterize rheological properties of lipid bilayer membranes
  where $R_0$ is vesicle radius $K_B$ is bending rigidity \citep{VLAHOVSKA2009775, keller_skalak_1982}. 
  Setting $\text{Ca}$ and $\tilde{\text{Ca}}$ equal, we can derive
  \[
  K_B = \frac{\mu \dot \gamma_* R_0^3}{\text{Ca}}
  \]
  which yields $K_B \approx 60$ pN $\text{nm}^{-1}$, which is a reasonable
  value for the bending modulus of a lipid bilayer, for example \citep{NaTr00}.}

@@

\subsection{\ReviewerOne{Stretching and Permeability}}
\label{sec:ves_in_quiescent}

@@

In membrane continuum  mechanics, changes in membrane surface area $A$ give rise to a tension
$\tilde \sigma = K_A(A/A_0 - 1)$ where $A_0$ is the reference surface area \cite{NaTr00}. 
For a two-dimensional membrane, the tension is 
\begin{align}
\label{eq:stretch}
\tilde \sigma = K_A\left(L/L_0 - 1 \right),
\end{align}
where $L$ and $L_0$ are arc length and resting length,
respectively. 
Membrane tension gives rise to 
a Laplace pressure 
\begin{align}
\label{eq:LaplacePressure}
\Delta p = \tilde \sigma/R,
\end{align}
where $\Delta p$ is the jump in far-field pressure to internal pressure and
$R = L/2\pi$.
In order to relieve Laplace pressure, 
water flows across the membrane. The rate of efflux is governed by  
\begin{align}
  \label{eq:perm} 
  \frac{dA}{dt} = - P\Delta p,
\end{align}
where $P = K/(\mu \Delta x)$ is a constant, $K$ is hydraulic permeability, 
$\mu$ is the solvent viscosity, and $\Delta x$ is the 
membrane thickness, about 5 nm
(\cite{chabanon2017, qua-gan-you2021}).

@@

Our current numerical method does not support a viscosity contrast
like that found in
\citep{PhysRevLett.106.158103,
  PhysRevE.75.016313,
  Kaoui2016}.
The JP vesicle is permeable and the mixture of two different
fluids requires an introduce of a contact line between
the two phases.

Our future goals include extending the current framework to a
three-dimensional JP vesicle system. This will require additional
algorithmic implementation including a fast summation method such as the
fast multipole method. Oother research directions include
fluctuating hydrodynamics for Brownian suspensions~(\cite{Bao2018}),
%introduce thermal fluctuations into the model. 
%This will result in Janus particles that include Brownian motion, 
critical to understanding membrane diffusion,
and an analysis of the relationship between particle shape on effective
properties.
Finally, more
physical boundary conditions for the HAP model will allow us to draw
comparisons between computational and laboratory experiments.
